---
title: "R Notebook"
output: html_notebook
---

#### This code is a snapshot of the complete code presented in the file "ranking.methods". It seeks to show more easily the split between training and testing datasets and how it is applied to the 2 types of approach: dynamic and static. 

### Load needed functions
```{r}
## CREATE DATAFRAME FUNCTION
create.dataframe<- function(training.data){
  
  List<- training.data[,3:4]
  Individuals<-c(as.character(List$from),as.character(List$to))
  
  #Delete replications
  Individual.vector<-unique(Individuals)
  nb.individual <- length(Individual.vector)
  
  #create a data frame out of a transposed vector
  Obs.Frame = as.data.frame(t(Individual.vector))
  #change the names of the dataframe to be IDs
  colnames(Obs.Frame) <- Individual.vector # The outcome isn't the best as the fiRBM row contains all the IDs again, but it's an easy fix (we will delete the fiRBM row later)
  
  return(Obs.Frame)
}

## EDGELIST FUNCTION
create.an.edgeList<-function(nn){
  
  Edgelist<-dplyr::count(nn, winner , loser)
  
  return(Edgelist)
}

rbind.all.columns <- function(x, y) {
  
  x.diff <- setdiff(colnames(x), colnames(y))
  y.diff <- setdiff(colnames(y), colnames(x))
  
  x[, c(as.character(y.diff))] <- NA
  
  y[, c(as.character(x.diff))] <- NA
  
  return(rbind(x, y))
}

```

### Import dataset. 
```{r}
## Import dominance data
Dominance <- read.csv ("dominance.data.csv")

##FYI: in the result column a win =1, a loss=2, a draw=3 and unknown result = 4
## Get rid of unknown outcomes.
Dominance.clear<-Dominance[-which(Dominance$result=="4"),]

## Get rid of individuals who appear in testing dataset only as we wont have any ranks associated to them from the training dataset.
Dominance.clear2<-Dominance.clear[-which (Dominance.clear$to=="macy"),]
Dominance.clear3<-Dominance.clear2[-which (Dominance.clear2$from=="macy"),]
Dominance.clear4<-Dominance.clear3[-which (Dominance.clear3$to=="rodr"),]
Dominance.clear5<-Dominance.clear4[-which (Dominance.clear4$from=="rodr"),]
Dominance.clear6<-Dominance.clear5[-which (Dominance.clear5$to=="balu"),]
Dominance.clear7<-Dominance.clear6[-which (Dominance.clear6$from=="balu"),]
Dominance.clear8<-Dominance.clear7[-which (Dominance.clear7$to=="nige"),]
## Set up date format and make sure interactions are chronologically ordered
Dominance.clear8$date <- lubridate::ymd(as.character(Dominance.clear8$date))
Dominance.df <- Dominance.clear8[order(Dominance.clear8$date),] 
```

###### PART 1 
#### Split the dataset into training and testing
```{r}
## here we want to keep the chronological order hence using the function filter and not just the sample function
# TRAINING dataset (80% OF THE DATA)
training.data<- Dominance.df %>% filter(date<="2017-04-25")

# TESTING dataset (20% OF THE DATA)
testing.data<- Dominance.df %>% filter(date>"2017-04-25")
```

#### Create main dataframe that will store the ranks
```{r}
## create main df to store ranks
main.df<- create.dataframe(training.data)
main.df<-cbind(main.df,Method = "method")
main.df$Method<-as.character(main.df$Method)
```

### First start by inferring and extracting the individual ranks, using the training dataset. We offer two examples: one using a dynamic approach (Elo-rating) and the second using a matrix-based approach (David's scores)

## ORIGINAL ELO RATING MEHTOD
```{r}
## Run elo on the TRAINING data ONLY
# upload presence grid
presence.grid.train.data<- read.csv("daily.presence.csv", header = T)
presence.grid.train.data$Date<- lubridate::ymd(as.character(presence.grid.train.data$Date))

# Check that the train data are doing ok format wise.
seqcheck(winner=as.character(training.data$winner), loser=as.character(training.data$loser), Date=training.data$date, draw = training.data$draw, presence = presence.grid.train.data)

# Run elo: draws AND presence grid are included. 
elo.scores <- elo.seq(winner=training.data$winner, loser=training.data$loser, Date=training.data$date,  runcheck=FALSE, draw =  training.data$draw, presence = presence.grid.train.data)

# Select the scores for the latest date in the training dataset
latest.scores<-extract_elo(elo.scores, "2017-04-25") # NB: individuals who died during this period will have a "NA" score

## Store elo rating per individual
elo.df<-as.data.frame(t(sort(latest.scores)))
elo.df[2,]<- c(41:1)
elo.df<-elo.df[-1,]
```

## DAVID'S SCORES using the compete package and the method "p"
```{r}
## Isolate winner-loser to get a matrix with extra column giving the results
outcome.data<- training.data[,c("winner","loser","result")]
outcome.data$result[outcome.data$result %in% "1"]<-"W"
outcome.data$result[outcome.data$result %in% "2"]<-"L"
outcome.data$result[outcome.data$result %in% "3"]<-"T"

matrix.result<- get_wl_matrix(outcome.data, ties = "keep")

## Method P
David.score.p<-ds(matrix.result, norm = TRUE, type = "P")

## Store david's scores
dataframe<-as.data.frame (t(sort(David.score.p)))
dataframe[2,]<-c(62:1)
dataframe<-dataframe[-1,]
```

#compile elo and david's scores dataframes
```{r}
main.df<-as.data.frame(t(rbind.fill(list(elo.df, dataframe))))
colnames(main.df)<- c("elo","david","ID")
main.df$elo<-as.integer(main.df$elo)
main.df$david<-as.integer(main.df$david)
```

## Create dataframe to store data when looking at whether ranks match aggressive outcomes in the testing dataset

```{r}
Decay.dataframe<- as.data.frame(matrix(0, ncol = 2, nrow = nrow(testing.data)))
colnames(Decay.dataframe)<- c("Original.elo", "compete.p")

## df to store day.nb
Day.df<- as.data.frame(matrix(0, ncol = 2, nrow = nrow(testing.data)))
colnames(Day.df)<- c("Original.elo", "compete.p")
```


## For loop to test methods' reliability: rank orders are used to assess how well they matched the outcome of dyadic aggressive interactions in the TESTING dataset (yes =1 or no=0). The proportion of reliably predicted outcomes is then translated into a percentage.

```{r}
main.df$Winner <- factor(main.df$ID, levels=levels(testing.data$winner))
main.df$Loser <- factor(main.df$ID, levels=levels(testing.data$loser))
main.df <- main.df [,c(3,4,5,1,2)]

result.efficiency<- vector(length = ncol(main.df))
is.rank.true<- vector(length = nrow(testing.data))
total <- nrow(testing.data)

## Add day - aggressor - victim for analysis purposes.
day.nb<- vector(length = nrow(testing.data))

# create progress bar
pb <- txtProgressBar(min = 0, max = total, style = 3)

for (j in 4:ncol(main.df)) {
  
  for(i in 1:total){
    
    winner_ID_i <- testing.data$winner[i]
    loser_ID_i <- testing.data$loser[i]
    match_winner <- which(unlist(main.df$Winner) == winner_ID_i)
    match_loser<- which (unlist(main.df$Loser)== loser_ID_i)
    
    if( main.df[match_winner,j]<=main.df[match_loser,j]){
      is.rank.true[i] <- 1
      day.nb[i]<- testing.data$day_nb[i]
      
    }else{
      is.rank.true[i] <- 0
      day.nb[i]<- testing.data$day_nb[i]
  
    }
    
    
    # Calculate efficiency
    efficiency.prediction <- sum(na.omit(is.rank.true))/length(na.omit(is.rank.true))
    percentage.efficiency<- efficiency.prediction*100
    
    Decay.dataframe[j-3]<-is.rank.true
    Day.df[j-3]<-day.nb
    
    # update progress bar
    setTxtProgressBar(pb, i)
  } 
  result.efficiency[j]<- percentage.efficiency
}
close(pb)   

# Store the result vector in dataframe
Efficiency.dataframe<- as.data.frame(matrix(0, ncol = 2, nrow = 1))
colnames(Efficiency.dataframe)<- c("Original.elo","compete.p")


## Store the tries...
Efficiency.dataframe$Original.elo<- result.efficiency[4]
Efficiency.dataframe$compete.p<- result.efficiency[5]
```

###### PART 2: modify the training dataset length and calculate ranks for each length of it

#####Isolate individual ID that we will bind to the matching ranks
```{r}
List<- training.data[,3:4]
Individuals<-c(as.character(List$from),as.character(List$to))

Individual.vector<-unique(Individuals)
df.ind<- data.frame(Individual.vector)
colnames(df.ind)<- c("ID")
```

### Run a loop to extract ranks for each different training dataset length. Loop done for each tested method.

## ELO-RATING ORIGINAL
```{r}
## For my own simplicity, in what follows i use the number of days (instead of the date column).
# Add day_nb column to the presence grid. 
presence.grid.train.data$Day<- as.numeric(presence.grid.train.data$Date) - 16437 +2
presence.grid.train.data<- presence.grid.train.data[, c(1,64,2:63)]

# Run loop

df.elo<- create.dataframe(training.data)
Day<-vector()

for (m in seq(6,786, by=60)) {
  
  windowStart<-m
  sub.df<-  training.data %>% filter(day_nb>=m)
  sub.grid<-  presence.grid.train.data[presence.grid.train.data[,2]>=m,]
  
  ## Run elo: avec draw AND presence included. 
  res<- elo.seq(winner=as.character(sub.df$winner), loser=as.character(sub.df$loser), Date=sub.df$date,runcheck=FALSE, draw =  sub.df$draw, presence = sub.grid)
  
  ## Select the scores for the latest date
  latest.scores<-extract_elo(res)
  
  ## Store elo ratings
  dataframe<-as.data.frame(t(sort(latest.scores)))
  dataframe[2,]<- c(length(latest.scores):1)
  dataframe<-dataframe[-1,]
  df.elo<-rbind.fill(list(df.elo, dataframe))
  Day[length(Day)+1]<-windowStart
  
}

#store ranks
df.elo<- df.elo[-1,]
Elo.ranks<- cbind(df.elo, Day)
Elo.ranks<- as.data.frame(t(Elo.ranks))
colnames(Elo.ranks)<-as.character(unlist(Elo.ranks[63,]))
Elo.ranks<-Elo.ranks[-63,]
Elo.ranks[,15]<- c("Elo.original")
names(Elo.ranks)[length(names(Elo.ranks))]<-"Method" 
Elo.ranks<-cbind(Elo.ranks,df.ind)
```

## DAVID'S SCORE
## Compete package
```{r}
Day<-vector()
main.df.p<- create.dataframe(training.data)

# Run loop
for (m in seq(6,786, by=60)) {
  
  windowStart<-m
  sub.df<-  training.data %>% filter(day_nb>=m)
  
  ## Isolate winner-loser to get a matrix with extra column giving the results
  outcome.data<- sub.df[,c("winner","loser","result")]
  outcome.data$result[outcome.data$result %in% "1"]<-"W"
  outcome.data$result[outcome.data$result %in% "2"]<-"L"
  outcome.data$result[outcome.data$result %in% "3"]<-"T"
  
  matrix<- get_wl_matrix(outcome.data, ties = "keep")
  
  ## Get rank (DS), Type P
  David.score.p<-ds(matrix, norm = TRUE, type = "P")
  dataframe<-as.data.frame(t(sort(David.score.p)))
  dataframe[2,]<- c(length(David.score.p):1)
  dataframe<-dataframe[-1,]
  main.df.p<-rbind.fill(list(main.df.p, dataframe))
  
  Day[length(Day)+1]<-windowStart
  
}

#Store ranks
main.df.p<- main.df.p[-1,]
DS.ranks.P<- cbind(main.df.p, Day)
DS.ranks.P<- as.data.frame(t(DS.ranks.P))
colnames(DS.ranks.P)<-as.character(unlist(DS.ranks.P[63,]))
DS.ranks.P<-DS.ranks.P[-63,]
DS.ranks.P[,15]<- c("DS.compete.P")
names(DS.ranks.P)[length(names(DS.ranks.P))]<-"Method" 

Ranks.compete.P<-cbind(DS.ranks.P,df.ind)
```

## compile both dataframes
```{r}
rank.training<- rbind(Ranks.compete.P,Elo.ranks)
```

## Prediction loop to be run in order to get decay for each method and training dataset length
# Create dataframe to store values
```{r}
prefix<- "month"
suffix<- seq(28,2,-2)

Decay.dataframe<- as.data.frame(matrix(0, ncol = 15, nrow = nrow(testing.data)))
colnames(Decay.dataframe)<- paste(prefix, suffix, sep="_")

## df to store day.nb
Day.df<-as.data.frame(matrix(0, ncol = 15, nrow = nrow(testing.data)))
colnames(Day.df)<- paste(prefix, suffix, sep="_")
```

## For loop to test methods's reliability: do ranks match aggressive outcomes in the testing dataset?
```{r}
#set up dataframe
rank.training$Winner <- factor(rank.training$ID, levels=levels(testing.data$winner))
rank.training$Loser <- factor(rank.training$ID, levels=levels(testing.data$loser))
rank.training <- rank.training[,c(15:18,1:14)]

Efficiency.dataframe<- as.data.frame(matrix(0, ncol = 14, nrow = 1))
colnames(Efficiency.dataframe)<-paste(prefix, suffix, sep="_")
result.efficiency<- vector(length = 14)
is.rank.true<- vector(length = nrow(testing.data))
total <- nrow(testing.data)
day.nb<- vector(length = nrow(testing.data))

# create progress bar
pb <- txtProgressBar(min = 0, max = total, style = 3)

#Run loop: here make k vary between 1 and length (unique(rank.training$Method)), here = 2  as only present 2 mwethods
k<-1
subset.data <-rank.training[rank.training$Method==(levels(as.factor(rank.training$Method))[k]),]

for ( j in 5:ncol(subset.data)) {
  
  subset.rank<-  subset.data[,c(3,4,j)]
  
  for(i in 1:total){
    
    winner_ID_i <- testing.data$winner[i]
    loser_ID_i <- testing.data$loser[i]
    match_winner <- which(unlist(subset.rank$Winner) == winner_ID_i)
    match_loser<- which (unlist(subset.rank$Loser)== loser_ID_i)
    
    if ( is.na(subset.rank[match_loser,3]) | is.na(subset.rank[match_winner,3]) ){
      is.rank.true[i] <- NA
      day.nb[i]<- NA
      
    } else if (as.numeric(as.character(subset.rank[match_winner,3])) <=as.numeric(as.character(subset.rank[match_loser,3]))){
      is.rank.true[i] <- 1
      day.nb[i]<- testing.data$day_nb[i]
      
    } else if(as.numeric(as.character(subset.rank[match_winner,3])) >=as.numeric(as.character(subset.rank[match_loser,3]))){
      is.rank.true[i] <- 0
      day.nb[i]<- testing.data$day_nb[i]
      
    }
  }
  
  # Calculate efficiency
  efficiency.prediction <- sum(na.omit(is.rank.true))/length(na.omit(is.rank.true))
  percentage.efficiency<- efficiency.prediction*100
  Decay.dataframe[j-4]<-is.rank.true
  Day.df[j-4]<-day.nb
  
  # update progress bar
  setTxtProgressBar(pb, i)
  result.efficiency[j-4]<- percentage.efficiency
}

Efficiency.dataframe$month_28<- result.efficiency[1]
Efficiency.dataframe$month_26<- result.efficiency[2]
Efficiency.dataframe$month_24<- result.efficiency[3]
Efficiency.dataframe$month_22<- result.efficiency[4]
Efficiency.dataframe$month_20<- result.efficiency[5]
Efficiency.dataframe$month_18<- result.efficiency[6]
Efficiency.dataframe$month_16<- result.efficiency[7]
Efficiency.dataframe$month_14<- result.efficiency[8]
Efficiency.dataframe$month_12<- result.efficiency[9]
Efficiency.dataframe$month_10<- result.efficiency[10]
Efficiency.dataframe$month_8<- result.efficiency[11]
Efficiency.dataframe$month_6<- result.efficiency[12]
Efficiency.dataframe$month_4<- result.efficiency[13]
Efficiency.dataframe$month_2<- result.efficiency[14]

close(pb)   

#Store diff dataframes
# for k<-1
Efficiency.dataframe.CompP<-Efficiency.dataframe
Efficiency.dataframe.CompP[,15]<-"Compete.P"

#for k<-2
Efficiency.dataframe.Elo <-Efficiency.dataframe
Efficiency.dataframe.Elo[,15]<-"Elo.original"
method.efficiency<- rbind.fill(Efficiency.dataframe.Elo,Efficiency.dataframe.CompP)
names(method.efficiency)[length(names(method.efficiency))]<-"Method" 
```

###### PART 3: keep the training dataset constant and modify the testing dataset length

####  New split of training and testing datasets. Here we split sooner so we have a much bigger testing dataset to play around. 
```{r}
## Split data
# TRAINING dataset
newtraining.data<- Dominance.df %>% filter(date<="2015-07-04")

# TESTING dataset
newtesting.data<- Dominance.df %>% filter(date>"2015-07-04")
```

## Obtain dataframe of individual IDs
```{r}
List<- newtraining.data[,3:4]
Individuals<-c(as.character(List$from),as.character(List$to))

Individual.vector<-unique(Individuals)
df.ind<- data.frame(Individual.vector)
colnames(df.ind)<- c("ID")
```

##Get  presence grid ready
```{r}
presence.grid.train.data<- read.csv("daily.presence.testing.variation.csv", header = T)
presence.grid.train.data$Date<- lubridate::ymd(as.character(presence.grid.train.data$Date))
```

### Extract ranks with each method using the new training dataset
## ORIGINAL ELO-RATING
```{r}
df.elo<- create.dataframe(newtraining.data)

## Run elo: avec draw AND presence included. 
seqcheck(winner=as.character(newtraining.data$winner), loser=as.character(newtraining.data$loser), Date=newtraining.data$date, draw = newtraining.data$draw, presence = presence.grid.train.data)

result.elo <- elo.seq(winner=as.character(newtraining.data$winner), loser=as.character(newtraining.data$loser), Date=newtraining.data$date,runcheck=FALSE, draw =  newtraining.data$draw, presence = presence.grid.train.data)

## Select the scores for the latest date
latest.scores<-extract_elo(result.elo)

## Store mother's elo ratings
## Store elo rating per individual
elo.df<-as.data.frame(t(sort(latest.scores)))
elo.df[2,]<- c(39:1)
elo.df<-elo.df[-1,]
```

## DAVID'S SCORES
#Compete package
```{r}
## Create storing df
main.df.p<- create.dataframe(newtraining.data)

## Isolate winner-loser to get a matrix with extra column giving the results
outcome.data<- newtraining.data[,c("winner","loser","result")]
outcome.data$result[outcome.data$result %in% "1"]<-"W"
outcome.data$result[outcome.data$result %in% "2"]<-"L"
outcome.data$result[outcome.data$result %in% "3"]<-"T"

matrix<- get_wl_matrix(outcome.data, ties = "keep")

## Get rank (DS), Type P
David.score.p<-ds(matrix, norm = TRUE, type = "P")
## Store david's scores
dataframe<-as.data.frame (t(sort(David.score.p)))
dataframe[2,]<-c(43:1)
dataframe<-dataframe[-1,]
dataframe[2,]<-names(dataframe)
```

#compile elo and david's scores dataframes
```{r}
rank.df<-as.data.frame(t(rbind.fill(list(elo.df, dataframe))))
colnames(rank.df)<- c("elo","david","ID")
rank.df$elo<-as.integer(as.character(rank.df$elo))
rank.df$david<-as.integer(as.character(rank.df$david))

rank.df<-melt(rank.df, id.vars = c("ID"))
colnames(rank.df)<- c("ID", "Method", "Ranks")
```

## For loop to test if the rank orders match with dyadic outcomes from testing dataset
```{r}
rank.df$Winner <- factor(rank.df$ID, levels=levels(newtesting.data$winner))
rank.df$Loser <- factor(rank.df$ID, levels=levels(newtesting.data$loser))
rank.df <- rank.df[,c(1,4,5,3,2)]

k<-2 ## manually make k vary between 1 and 2 (total number of methods tested)
subset.data <-rank.df[rank.df$Method==(levels(as.factor(rank.df$Method))[k]),]

Efficiency.dataframe<- as.data.frame(matrix(0, ncol = 15, nrow = 1))
colnames(Efficiency.dataframe)<-paste(prefix, suffix, sep="_")
result.efficiency<- vector()

# create progress bar
pb <- txtProgressBar(min = 0, max = total, style = 3)


start <- 186
end<-1086
windowsize<- 60
j<-1
while (start + windowsize<=end) {
  
  m<-start + windowsize
  sub.prediction<-  newtesting.data %>% filter(day_nb<=m)
  total<- nrow(sub.prediction)
  is.rank.true<- vector(length = nrow(sub.prediction))
  
  ## Add day for analysis purposes.
  day.nb<- vector(length = nrow(sub.prediction))
  original<- vector(length = nrow(newtesting.data))
  
  max.len.decay = max(length(is.rank.true), length(original))
  max.len.day = max(length(day.nb), length(original))
  
  for(i in 1:total){
 
    winner_ID_i <- sub.prediction$winner[i]
    loser_ID_i <- sub.prediction$loser[i]
    match_winner <- which(unlist(subset.data$Winner) == winner_ID_i)
    match_loser<- which (unlist(subset.data$Loser)== loser_ID_i)
    

    if ( length(subset.data[match_loser,4])==0 | length(subset.data[match_winner,4])==0 ){
      is.rank.true[i] <- NA
      day.nb[i]<- NA
    
    } else if (as.numeric(as.character(subset.data[match_winner,4])) <=as.numeric(as.character(subset.data[match_loser,4]))){
      is.rank.true[i] <- 1
      day.nb[i]<- sub.prediction$day_nb[i]
      
    } else if(as.numeric(as.character(subset.data[match_winner,4])) >=as.numeric(as.character(subset.data[match_loser,4]))){
      is.rank.true[i] <- 0
      day.nb[i]<- sub.prediction$day_nb[i]
    }
    
    # Calculate efficiency
    efficiency.prediction <- sum(na.omit(is.rank.true))/length(na.omit(is.rank.true))
    percentage.efficiency<- efficiency.prediction*100
    is.rank.true = c(is.rank.true, rep(NA, max.len.decay - length(is.rank.true)))
    Decay.dataframe[j]<-is.rank.true
    day.nb=c(day.nb, rep(NA,max.len.day-length(day.nb)))
    Day.df[j]<-day.nb
    
    # update progress bar
    setTxtProgressBar(pb, i)
    result.efficiency[j]<- percentage.efficiency
    
  } 
  
  windowsize<- windowsize+60
  j<-j+1

}

Efficiency.dataframe$month_2<- result.efficiency[1]
Efficiency.dataframe$month_4<- result.efficiency[2]
Efficiency.dataframe$month_6<- result.efficiency[3]
Efficiency.dataframe$month_8<- result.efficiency[4]
Efficiency.dataframe$month_10<- result.efficiency[5]
Efficiency.dataframe$month_12<- result.efficiency[6]
Efficiency.dataframe$month_14<- result.efficiency[7]
Efficiency.dataframe$month_16<- result.efficiency[8]
Efficiency.dataframe$month_18<- result.efficiency[9]
Efficiency.dataframe$month_20<- result.efficiency[10]
Efficiency.dataframe$month_22<- result.efficiency[11]
Efficiency.dataframe$month_24<- result.efficiency[12]
Efficiency.dataframe$month_26<- result.efficiency[13]
Efficiency.dataframe$month_28<- result.efficiency[14]
Efficiency.dataframe$month_30<- result.efficiency[15]

close(pb)   

#Store diff dataframes depending on k
Efficiency.dataframe.CompP<-Efficiency.dataframe
Efficiency.dataframe.CompP[,16]<-"Compete.P"
Efficiency.dataframe.Elo <-Efficiency.dataframe
Efficiency.dataframe.Elo[,16]<-"Elo.original"


method.efficiency<- rbind.fill(Efficiency.dataframe.Elo,Efficiency.dataframe.CompP)
names(method.efficiency)[length(names(method.efficiency))]<-"Method" 
```

